---
title: "California Housing"
output: html_document
editor_options: 
  chunk_output_type: console
---

<style type="text/css">
.title {
display: none;
}

</style>

<div class="row" style="padding-top: 0px;"> 
<div class="col-sm-6">

```{r Load libraries, echo = FALSE, warning = FALSE, message = FALSE}
setwd("C:/Users/zaliv/OneDrive/Desktop/Data Projects/CaliforniaHousingProject")

library(tidyverse)
library(tidycensus)
library(ggplot2)
library(spData)
library(sf)
library(rcartocolor)
library(cowplot)
library(tigris)
library(rgdal)    # for readOGR and others
library(sp)       # for spatial objects
library(leaflet)  # for interactive maps (NOT leafletR here)
library(maptools)
library(PerformanceAnalytics)
library(data.table)
library(hrbrthemes)
library(leaflet.extras)

options(tigris_use_cache = TRUE)
```
```{r Load data, echo = FALSE, warning = FALSE, message = FALSE}
CAPermits <- read.csv('CaliforniaPermits.csv') %>%
  mutate(NumUnits = as.numeric(gsub(",","",Total.Units)))

CALabor <- read.csv('Local_Area_Unemployment_Statistics__LAUS_.csv') %>%
  mutate(
    AreaType = ï..Area.Type,
    SeasonallyAdjusted = Seasonally.Adjusted..Y.N.)

AllACSVariables <- load_variables(year = 2019, 
                                  dataset = "acs5", 
                                  cache = TRUE)

get_acs_data <- function(ACSYears, variable) {
  
  ACSData <- map_dfr(
    ACSYears, 
    ~get_acs(
      geography = "county",
      variables = variable,
      year = .x,
      state = "CA",
      survey = "acs5",
      geometry = FALSE
    ),
    .id = "year") %>%
    mutate(County.Name = str_sub(NAME,end = -13))
  
  ACSData <- subset(ACSData, select = -c(moe))
}

ACSYears <- lst(2009, 2019)

CAPOPULATION <- "B01003_001"
CARENTLOW <- "B25057_001"
CARENTMED <- "B25058_001"
CARENTHIGH <- "B25059_001"
CAHOUSELOW <- "B25076_001"
CAHOUSEMED <- "B25077_001"
CAHOUSEHIGH <- "B25078_001"
CAOWNERWHITE <- "B25003A_002"
CAOWNERBLACK <- "B25003B_002"
CAOWNERLATINO <- "B25003I_002"


CAPopulation <- get_acs_data(ACSYears, CAPOPULATION)
CARentLow <- get_acs_data(ACSYears, CARENTLOW)
CARentMed <- get_acs_data(ACSYears, CARENTMED)
CARentHigh <- get_acs_data(ACSYears, CARENTHIGH)
CAHouseLow <- get_acs_data(ACSYears, CAHOUSELOW)
CAHouseMed <- get_acs_data(ACSYears, CAHOUSEMED)
CAHouseHigh <- get_acs_data(ACSYears, CAHOUSEHIGH)
CAOwnerWhite <- get_acs_data(ACSYears, CAOWNERWHITE)
CAOwnerBlack <- get_acs_data(ACSYears, CAOWNERBLACK)
CAOwnerLatino <- get_acs_data(ACSYears, CAOWNERLATINO)

```
```{r Clean Data, echo = FALSE, warning = FALSE, message = FALSE}
CleanCAPermits <- CAPermits %>% 
  filter(Survey.Date >= 2009 & Survey.Date <= 2019)

CleanCALabor <- CALabor %>% 
  filter(AreaType == 'County') %>%
  filter(Year == 2009 | Year == 2019) %>%
  filter(Month == 'January') %>%
  filter(SeasonallyAdjusted == 'N') %>%
  mutate(County.Name = Area.Name,
         Labor.Force = as.numeric(gsub(",","",Labor.Force))) %>%
  subset(select = c(County.Name, Year, Labor.Force))

CleanCAPopulation <- CAPopulation
CleanCARentLow <- CARentLow
CleanCARentMed <- CARentMed
CleanCARentHigh <- CARentHigh

CleanCAHouseLow <- CAHouseLow
CleanCAHouseMed <- CAHouseMed
CleanCAHouseHigh <- CAHouseHigh

CleanCAOwnerWhite <- CAOwnerWhite
CleanCAOwnerBlack <- CAOwnerBlack
CleanCAOwnerLatino <- CAOwnerLatino


```
```{r Spread Data, echo = FALSE, warning = FALSE, message = FALSE}
SpreadCAPermits <- CleanCAPermits
SpreadCALabor <- CleanCALabor  %>% spread(Year, Labor.Force)

spread_data <- function(DataTable) {SpreadDataTable <- DataTable %>% spread(year, estimate)}

SpreadCAPopulation <- spread_data(CleanCAPopulation)
SpreadCARentLow <- spread_data(CleanCARentLow)
SpreadCARentMed <- spread_data(CleanCARentMed)
SpreadCARentHigh <- spread_data(CleanCARentHigh)
SpreadCAHouseLow <- spread_data(CleanCAHouseLow)
SpreadCAHouseMed <- spread_data(CleanCAHouseMed)
SpreadCAHouseHigh <- spread_data(CleanCAHouseHigh)
SpreadCAOwnerWhite <- spread_data(CleanCAOwnerWhite)
SpreadCAOwnerBlack <- spread_data(CleanCAOwnerBlack)
SpreadCAOwnerLatino <- spread_data(CleanCAOwnerLatino)
```
```{r Chunk 2, echo = FALSE, warning = FALSE, message = FALSE}

calculate_growth_rate <- function(DataTable) {GrowthRate <- (DataTable$"2019" - DataTable$"2009")/
  DataTable$"2009"}
calculate_change <- function(DataTable) {LaborChange <- DataTable$"2019" - DataTable$"2009"}

SpreadCALabor <- SpreadCALabor %>%
  mutate(
    LaborGrowthRate = calculate_growth_rate(SpreadCALabor),
    LaborChange = calculate_change(SpreadCALabor)
  )

SpreadCAPopulation <- SpreadCAPopulation %>%
  mutate(
    PopGrowthRate = calculate_growth_rate(SpreadCAPopulation),
    PopChange = calculate_change(SpreadCAPopulation)
  )

SpreadCARentLow <- SpreadCARentLow %>%
  mutate(
    LowRentGrowthRate = calculate_growth_rate(SpreadCARentLow),
    LowRentChange = calculate_change(SpreadCARentLow)
  )

SpreadCARentMed <- SpreadCARentMed %>%
  mutate(
    MedRentGrowthRate = calculate_growth_rate(SpreadCARentMed),
    MedRentChange = calculate_change(SpreadCARentMed)
  )

SpreadCARentHigh <- SpreadCARentHigh %>%
  mutate(
    HighRentGrowthRate = calculate_growth_rate(SpreadCARentHigh),
    HighRentChange = calculate_change(SpreadCARentHigh)
  )

SpreadCAHouseLow <- SpreadCAHouseLow %>%
  mutate(
    LowHouseGrowthRate = calculate_growth_rate(SpreadCAHouseLow),
    LowHouseChange = calculate_change(SpreadCAHouseLow)
  )

SpreadCAHouseMed <- SpreadCAHouseMed %>%
  mutate(
    MedHouseGrowthRate = calculate_growth_rate(SpreadCAHouseMed),
    MedHouseChange = calculate_change(SpreadCAHouseMed)
  )

SpreadCAHouseHigh <- SpreadCAHouseHigh %>%
  mutate(
    HighHouseGrowthRate = calculate_growth_rate(SpreadCAHouseHigh),
    HighHouseChange = calculate_change(SpreadCAHouseHigh)
  )

SpreadCAOwnerWhite <- SpreadCAOwnerWhite %>%
  mutate(
    WhiteOwnerGrowthRate = calculate_growth_rate(SpreadCAOwnerWhite),
    WhiteOwnerChange = calculate_change(SpreadCAOwnerWhite)
  )

SpreadCAOwnerBlack <- SpreadCAOwnerBlack %>%
  mutate(
    BlackOwnerGrowthRate = calculate_growth_rate(SpreadCAOwnerBlack),
    BlackOwnerChange = calculate_change(SpreadCAOwnerBlack)
  )

SpreadCAOwnerLatino <- SpreadCAOwnerLatino %>%
  mutate(
    LatinoOwnerGrowthRate = calculate_growth_rate(SpreadCAOwnerLatino),
    LatinoOwnerChange = calculate_change(SpreadCAOwnerLatino)
  )

```
```{r Summarize, echo = FALSE, warning = FALSE, message = FALSE}

SummaryCAPermits <- SpreadCAPermits %>%
  group_by(County.Name) %>%
  summarize(NumPermits = sum(NumUnits))

SummaryCALabor <- SpreadCALabor %>% subset(select = c(County.Name, LaborGrowthRate, LaborChange))

SummaryCAPopulation <- SpreadCAPopulation %>% subset(select = c(County.Name, PopGrowthRate, PopChange))

SummaryCARentLow <- SpreadCARentLow %>% subset(select = c(County.Name, LowRentGrowthRate, LowRentChange))

SummaryCARentMed <- SpreadCARentMed %>% subset(select = c(County.Name, MedRentGrowthRate, MedRentChange))

SummaryCARentHigh <- SpreadCARentHigh %>% subset(select = c(County.Name, HighRentGrowthRate, HighRentChange))

SummaryCAHouseLow <- SpreadCAHouseLow %>% subset(select = c(County.Name, LowHouseGrowthRate, LowHouseChange))

SummaryCAHouseMed <- SpreadCAHouseMed %>% subset(select = c(County.Name, MedHouseGrowthRate, MedHouseChange))

SummaryCAHouseHigh <- SpreadCAHouseHigh %>% subset(select = c(County.Name, HighHouseGrowthRate, HighHouseChange))

SummaryCAOwnerWhite <- SpreadCAOwnerWhite %>% subset(select = c(County.Name, WhiteOwnerGrowthRate, WhiteOwnerChange))

SummaryCAOwnerBlack <- SpreadCAOwnerBlack %>% subset(select = c(County.Name, BlackOwnerGrowthRate, BlackOwnerChange))

SummaryCAOwnerLatino<- SpreadCAOwnerLatino %>% subset(select = c(County.Name, LatinoOwnerGrowthRate, LatinoOwnerChange))

```
```{r Combine Tables, echo = FALSE, warning = FALSE, message = FALSE}
# Need to figure out better way to do this

MergeCA1 <- merge(SummaryCAPermits, SummaryCALabor)
MergeCA2 <- merge(SummaryCAPopulation, SummaryCARentMed)
MergeCA3 <- merge(SummaryCARentHigh, SummaryCARentLow)
MergeCA4 <- merge(SummaryCAHouseLow, SummaryCAHouseMed)
MergeCA5 <- merge(SummaryCAHouseHigh, MergeCA4)
MergeCA6 <- merge(MergeCA1, MergeCA2)
MergeCA7 <- merge(MergeCA6, MergeCA3)
MergeCA8 <- merge(SummaryCAOwnerWhite, SummaryCAOwnerBlack)
MergeCA9 <- merge(MergeCA8, SummaryCAOwnerLatino)
MergeCA10 <- merge(MergeCA9, MergeCA5)

DataTable <- merge(MergeCA7, MergeCA10)

```
```{r Second Calculations, echo = FALSE, warning = FALSE, message = FALSE}

DataTable <- DataTable %>%
  mutate(LaborRatio = LaborChange/NumPermits,
         PopRatio = PopChange/NumPermits)

```
```{r Correlation Matrix, echo = FALSE, warning = FALSE, message = FALSE, include = FALSE}

CorrelationMatrix <- DataTable %>% subset(select = -c(County.Name))

chart.Correlation(CorrelationMatrix, histogram = TRUE)

plot(CorrelationMatrix$LaborRatio, CorrelationMatrix$RentChange)

```
```{r Chunk 9, echo = FALSE, warning = FALSE, message = FALSE}
us_states <- states(cb = FALSE, class = "sf") 

ca_state <- us_states %>% filter(STUSPS == "CA")
us_states <- 
  subset(us_states, 
         !NAME %in% c(
           "United States Virgin Islands",
           "Commonwealth of the Northern Mariana Islands",
           "Guam",
           "American Samoa",
           "Puerto Rico",
           "Alaska",
           "Hawaii",
           "California"
         ))


ca_counties <- counties("CA")

ca_county_names <- list_counties("CA")
ca_water <- area_water("CA", ca_county_names)

counties_for_water <- c("San Francisco County", "San Mateo County", 
                        "Alameda County", "Contra Costa County", 
                        "Sonoma County", 
                        "Santa Cruz County","Monterey County")
ca_water_2 <- area_water("CA",counties_for_water)


ca_water_na <- ca_water %>% drop_na(FULLNAME) %>% filter(AWATER > 10000000)

ca_places <- county_subdivisions(state = c("CA"))
nv_places <- county_subdivisions(state = c("NV"))
or_places <- county_subdivisions(state = c("OR"))
ut_places <- county_subdivisions(state = c("UT"))
az_places <- county_subdivisions(state = c("AZ"))
id_places <- county_subdivisions(state = c("ID"))

county_subdivisions_map <- rbind(ca_places, nv_places, or_places, ut_places,az_places,id_places)

ca_counties <- st_transform(ca_counties, crs = 4326) #, crs = 4326
us_states <- st_transform(us_states, crs = 4326) # , crs = 4326
county_subdivisions_map <- st_transform(county_subdivisions_map, crs = 4326)

ca_counties_bb <- st_as_sfc(st_bbox(ca_counties))
ca_counties_bb <- st_buffer(ca_counties_bb, dist = 10000)

ca_counties$values <- runif(nrow(ca_counties))

ca_counties_for_merge <- ca_counties %>% mutate(County.Name = NAMELSAD)
```

```{r Merge Map and Data, echo = FALSE, warning = FALSE, message = FALSE}

MapDataTable <- merge(DataTable, ca_counties_for_merge)

MapDataTable <- MapDataTable %>% mutate(LaborRatioSimple = 
                                          case_when(
                                            LaborRatio < 0 ~ '-2',
                                            LaborRatio <= "1" & LaborRatio >= "0" ~ '-1',
                                            LaborRatio >= "4" ~ '4',
                                            TRUE ~ as.character(LaborRatio)
                                          ),
                                        LaborRatioRounded = as.factor(floor(as.numeric(LaborRatioSimple))))


```

```{r Plot Variables, echo = FALSE, warning = FALSE, message = FALSE, include = FALSE}
reds <- tmaptools::get_brewer_pal("RdGy", 9)
cols <- c(reds[8], reds[7], reds[4], reds[3], reds[2], reds[1])

RentChange.labs <- c("Lower Quartile",
                     "Median Quartile",
                     "Upper Quartile")
names(RentChange.labs) <- c("LowRentChange",
                            "MedRentChange",
                            "HighRentChange")

HouseChange.labs <- c("Lower Quartile",
                      "Median Quartile",
                      "Upper Quartile")
names(HouseChange.labs) <- c("LowHouseChange",
                             "MedHouseChange",
                             "HighHouseChange")

OwnerGrowth.labs <- c("White",
                      "Black",
                      "Latino")
names(OwnerGrowth.labs) <- c("WhiteOwnerGrowthRate",
                             "BlackOwnerGrowthRate",
                             "LatinoOwnerGrowthRate")

col1 <- "#AAAAAA" 
col2 <- "#C7433F"

labels <- c("Lost Workers",
            "\u22641:1",
            ">1:1",
            ">2:1",
            ">3:1",
            ">4:1")


factorPal <- colorFactor(cols, MapDataTable$LaborRatioRounded)
```

# Map: Labor Force Size to Number of Permits by County

Shows CA county subdivisions with a ratio for change in labor force size to number of county permits issued from 2009-2019. Labor force size includes employed and unemployed people.

```{r Labor Force to Permits Interactive Map, echo = FALSE, include = TRUE, message = FALSE, warning = FALSE}

leaflet() %>%
  addPolygons(data = MapDataTable$geometry,
              fillColor = factorPal(MapDataTable$LaborRatioRounded),
              color = FALSE, # you need to use hex colors
              fillOpacity = 1,
              weight = 0.4,
              popup = ifelse(MapDataTable$LaborRatio >= 0,
                             paste0(MapDataTable$County.Name, " gained ", round(MapDataTable$LaborRatio,2)," workers for every housing permit issued from 2009 to 2019.","<br/>","<br/>",
                                    "Workers Gained: ", MapDataTable$LaborChange, "<br/>",
                                    "Permits Issued: ", MapDataTable$NumPermits),
                             paste0(MapDataTable$County.Name, " lost ", abs(round(MapDataTable$LaborRatio,2))," workers for every housing permit issued from 2009 to 2019.","<br/>","<br/>",
                                    "Workers Lost: ", abs(MapDataTable$LaborChange), "<br/>",
                                    "Permits Issued: ", MapDataTable$NumPermits)
              )
  ) %>%
  leaflet::addLegend(position = "bottomleft",
                     pal = factorPal,
                     values = MapDataTable$LaborRatioRounded,
                     title = "Workers:Permits",
                     opacity = 1,
                     labFormat = function(type, cuts, p) {
                       paste0(labels)
                     }) %>%
  addPolygons(data = ca_places$geometry,
              fill = FALSE,
              color = "#000000", 
              fillOpacity = 1,
              weight = 0.8) %>% 
  addPolygons(data = ca_water_na$geometry,
              fillColor = "#000000",
              color = FALSE, # you need to use hex colors
              fillOpacity = 1,
              weight = 2) %>%
  setMapWidgetStyle(list(background= "black")) 
```

```{r Labor Force Map, echo = FALSE, warning = FALSE, message = FALSE, fig.width=5, fig.height=11, include = FALSE}
MapDataTableForLabels <- MapDataTable %>% filter(LaborRatio >= 2.5) %>%
  mutate(County = str_sub(County.Name,end = -7))

ggplot() + 
  geom_sf(data = us_states, fill = "grey20", color = NA, size = 2) +
  geom_sf(data = MapDataTable, 
          aes(fill = LaborRatioRounded, 
              geometry = geometry), color = NA) +
  geom_sf(data = county_subdivisions_map, 
          aes(geometry = geometry), fill = NA, color = "white") + #grey10
  geom_sf(data = ca_water_2, 
          aes(geometry = geometry), fill = "white", color = NA) +
  geom_sf(data = ca_state, 
          aes(geometry = geometry), fill = NA, color = "white", size = .75) +
  geom_sf(data = us_states, fill = NA, color = 'black', size = .9) +
  labs(fill = "Workers:Permits",
       caption = "Time: 2009-2019") +
  scale_fill_manual(breaks = c("-2","-1","1","2","3","4"),
                    labels = c("Lost Workers",
                               "\u22641:1",
                               ">1:1",
                               ">2:1",
                               ">3:1",
                               ">4:1"),
                    values = cols) +
  theme_void() +
  theme(legend.title = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 12),
        legend.position = c(.17, .12),
        legend.direction = "vertical",
        legend.key.width = unit(10, "mm"),
        plot.background = element_rect(fill = NULL)) +
  coord_sf(xlim = st_bbox(ca_counties_bb)[c(1, 3)],
           ylim = st_bbox(ca_counties_bb)[c(2, 4)])


```

# Change in Rent by Permits Issued

Shows rent change for counties grouped by Workers:Permits < 1:1 (permits outnumbered workers) and Workers:Permits > 1:1 (workers outnumbered permits). On average, counties where workers outnumbered permits saw larger rent increases from 2009-2019.

```{r Workers Permits and Rent, echo = FALSE, warning = FALSE, message = FALSE, fig.width=7, fig.height=7}

FilteredDataTable <- DataTable %>% 
  filter(LaborRatio >= 0) %>% 
  mutate(TileSplit = ifelse(LaborRatio > 1, 2, 1))

RentChanges <- subset(FilteredDataTable, select = c(County.Name, 
                                                    LaborRatio,
                                                    LowRentChange,
                                                    MedRentChange,
                                                    HighRentChange,
                                                    TileSplit))

RentChangesMelt <- RentChanges %>% 
  reshape2::melt(id.vars = c("County.Name","LaborRatio","TileSplit"),
                 measure.vars = c("LowRentChange",
                                  "MedRentChange",
                                  "HighRentChange"))

RentChangesSummary <- RentChangesMelt %>%
  group_by(TileSplit, variable) %>%
  summarize(AvgChange = mean(value))

RentPlot <- ggplot(RentChangesSummary, aes(x = as.factor(TileSplit), y = AvgChange)) +
  geom_bar(aes(fill = as.factor(TileSplit)), 
           stat = 'identity', 
           color = 'black',
           size = 1) +
  geom_point(data = RentChangesMelt, aes(x = TileSplit, y = value), 
             alpha=.75, 
             color = 'black') +
  labs(title = "Change in Rent by County",
       subtitle = "Time: 2009-2019",
       x = NULL,
       y = "Change") + 
  scale_fill_manual(name = NULL,
                    breaks = c("1","2"),
                    labels = c("Workers < Permits", "Workers > Permits"),
                    values = c(col1, col2)) +
  scale_x_discrete(breaks=c("1","2"),
                   labels=NULL) +
  scale_y_continuous(labels=scales::dollar_format()) +
  facet_wrap(~variable, labeller = labeller(variable = RentChange.labs)) +
  theme_ipsum(grid="Y") +
  theme(legend.position = "bottom")

RentPlot

```

# Change in Home Value by Permits Issued

Shows home value change for counties grouped by Workers:Permits < 1:1 (permits outnumbered workers) and Workers:Permits > 1:1 (workers outnumbered permits). On average, counties where workers outnumbered permits saw larger increases in home value from 2009-2019.

```{r Workers Permits and House, echo = FALSE, warning = FALSE, message = FALSE, fig.width=7, fig.height=7}

HouseChanges <- subset(FilteredDataTable, select = c(County.Name, 
                                                     LaborRatio,
                                                     LowHouseChange,
                                                     MedHouseChange,
                                                     HighHouseChange,
                                                     TileSplit))

HouseChangesMelt <- HouseChanges %>% 
  reshape2::melt(id.vars = c("County.Name","LaborRatio","TileSplit"),
                 measure.vars = c("LowHouseChange",
                                  "MedHouseChange",
                                  "HighHouseChange"))

HouseChangesSummary <- HouseChangesMelt %>%
  group_by(TileSplit, variable) %>%
  summarize(AvgChange = mean(value))

HousePlot <- ggplot(HouseChangesSummary, aes(x = as.factor(TileSplit), y = AvgChange)) +
  geom_bar(aes(fill = as.factor(TileSplit)), 
           stat = 'identity', 
           color = 'black',
           size = 1) +
  geom_point(data = HouseChangesMelt, aes(x = TileSplit, y = value), 
             alpha=.75, 
             color = 'black') +
  labs(title = "Change in Home Value by County",
       subtitle = "Time: 2009-2019",
       x = NULL,
       y = "Change") + 
  scale_fill_manual(name = NULL,
                    breaks = c("1","2"),
                    labels = c("Workers < Permits", "Workers > Permits"),
                    values = c(col1, col2)) +
  scale_x_discrete(breaks=c("1","2"),
                   labels=NULL) +
  scale_y_continuous(labels=scales::dollar_format()) +
  facet_wrap(~variable, labeller = labeller(variable = HouseChange.labs)) +
  theme_ipsum(grid="Y") +
  theme(legend.position = "bottom")

HousePlot
```
```{r Combined Rent and House Plot,echo = FALSE, message=FALSE,warning = FALSE, fig.height=9, fig.width = 7, include = FALSE}

plot_grid(RentPlot + 
            theme(legend.position = 'none',
                  axis.title.y = element_blank()),
          HousePlot + 
            theme(axis.title.y = element_blank()), 
          cols = 1)

```

# Growth in Home Owners by Permits Issued

Shows home owner growth for counties grouped by Workers:Permits < 1:1 (permits outnumbered workers) and Workers:Permits > 1:1 (workers outnumbered permits). On average, counties where workers outnumbered permits saw slower growth in black and Latino home owners from 2009-2019.

```{r Home Owners, echo = FALSE, warning = FALSE, message = FALSE, fig.width=7, fig.height=7}

OwnerChanges <- subset(FilteredDataTable, select = c(County.Name, 
                                                     LaborRatio,
                                                     WhiteOwnerGrowthRate,
                                                     BlackOwnerGrowthRate,
                                                     LatinoOwnerGrowthRate,
                                                     TileSplit))

OwnerChangesMelt <- OwnerChanges %>% 
  reshape2::melt(id.vars = c("County.Name","LaborRatio","TileSplit"),
                 measure.vars = c("WhiteOwnerGrowthRate",
                                  "BlackOwnerGrowthRate",
                                  "LatinoOwnerGrowthRate"))

OwnerChangesSummary <- OwnerChangesMelt %>%
  filter(County.Name != 'Alpine County') %>%
  group_by(TileSplit, variable) %>%
  summarize(AvgChange = mean(value))

ggplot(OwnerChangesSummary, aes(x = as.factor(TileSplit), y = AvgChange)) +
  geom_bar(aes(fill = as.factor(TileSplit)), 
           stat = 'identity', 
           color = 'black',
           size = 1) +
  geom_point(data = OwnerChangesMelt, aes(x = TileSplit, y = value), 
             alpha=.75, 
             color = 'black') +
  labs(title = "Growth in Home Owners by County",
       subtitle = "Time: 2009-2019",
       x = NULL,
       y = "Growth") + 
  scale_fill_manual(name = NULL,
                    breaks = c("1","2"),
                    labels = c("Workers < Permits", "Workers > Permits"),
                    values = c(col1, col2)) +
  scale_x_discrete(breaks=c("1","2"),
                   labels=NULL) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  facet_wrap(~variable, labeller = labeller(variable = OwnerGrowth.labs)) +
  theme_ipsum(grid="Y") +
  theme(legend.position = "bottom")

```



</a>
</div>


</a>
</div>
